\appendix
\section*{Appendix: Signal Framework Details}
\label{sec:appendix}

This appendix provides complete documentation of the signal framework used for ETF selection. The framework consists of three stages: signal base computation, smoothing filter application, and indicator transformation.

\subsection*{Signal Bases}

Tables~\ref{tab:signal-bases} and \ref{tab:signal-bases-2} list all 95 signal bases computed from raw price data. Each signal captures a different aspect of ETF behavior relative to the core benchmark (ACWI). Signals are organized into categories: returns and alpha, relative strength, risk metrics, risk-adjusted performance, trend indicators, mean reversion signals, higher moments, correlation, momentum dynamics, signal disagreement, regime-adaptive signals, and seasonality.

\input{tables/signal_bases}

\subsection*{Smoothing Filters}

Table~\ref{tab:filters} describes the 14 causal smoothing filter types (24 configurations with different parameters) applied to each signal base. All filters are strictly causal---they only use past data at each time step, ensuring no look-ahead bias in the backtest. The filters range from simple exponential moving averages to adaptive filters (KAMA, regime-switching) and sophisticated signal processing methods (Kalman, Butterworth, Savitzky-Golay).

\input{tables/filters}

\subsection*{Indicator Transformations}

Table~\ref{tab:indicators} documents the 25 indicator transformations applied to filtered signals. These indicators capture different temporal and cross-sectional aspects of signal behavior, including momentum at multiple horizons, z-scores, trend ratios, mean reversion, velocity, acceleration, and volatility ratios.

\input{tables/indicators}

\subsection*{Feature Space Summary}

Combining all components yields the complete feature space:

\begin{itemize}
    \item \textbf{Signal Bases}: 95 fundamental signals
    \item \textbf{Smoothing Filters}: 24 causal filter configurations (including raw)
    \item \textbf{Indicators}: 25 transformations
    \item \textbf{Total Features}: $95 \times 24 \times 25 = 57{,}000$
\end{itemize}

Each feature is named using the convention: \texttt{signal\_\_filter\_\_indicator}, for example \texttt{vol\_boosted\_reversion\_\_hull\_63d\_\_xs\_rank}.

The GPU-accelerated computation pipeline processes all 57,000 features efficiently using:
\begin{itemize}
    \item CuPy for parallel computation on NVIDIA GPUs
    \item Streaming mode to limit memory usage
    \item Disk caching for rolling statistics to avoid redundant computation
    \item Vectorized operations for signal bases and indicators
\end{itemize}
